### 2009-09-16 latex/paper1/contents.tex
paper

One of the challenges of software engineering is the reduction of the time
needed to develop or maintain a software system. In

\cite{Brooks1995}

The development and maintenance of large software systems is a big challenge for
software engineering \cite{Northrop2006}. In order to reduce the development
time of such a system, a tempting idea is to assign more developers for the
project.

Because of communication costs between
developers, adding more developers to a project

Adding more developers to a software
project does not

A key to success is the ability to decompose a system into weakly-coupled
modules, so each module can be independently maintained by a distinct team.
Failing to do so results in duplicated code, non-parallelism, one's work
impacting another's work etc.

The ability do modularize depends decisively on a vast knowledge about the
system, how its different parts interact to accomplish the system's goal.

Unfortunately, in the case of legacy systems, such knowledge isn't available.
Depending on its size, it might take months to understand the system so well as
to find a good modularization. XXX \cite{Parnas1972}

POR ISSO SURGIRAM software clustering algorithms, also known as software
architecture recovery algorithms \cite{Pollet2007}. In its most common flavor,
these algorithms analyze the dependencies between implementation components,
such as classes, and then group them into modules such as there are few
dependencies between classes in distinct modules.

TODO: sw clustering is a task of reverse engineering \cite{Tonella2007}.

Software modularization recovery algorithms can, therefore, do in minutes what a
person would spend weeks or months. The question is: are the found
modularizations good? Are they similar to what a person would find? To answer
this question it's essencial to perform empirical evaluations envolving systems
with known reference modularizations.

The empirical evaluations consist of selecting a collection of systems with
known reference modularizations and then applying the algorithms to the systems.
The modularizations found by the algorithms are then compared to the reference
decompositions by a metric such as MoJo \cite{Tzerpos1999} or PrecisionRecall.

  \begin{figure*}[!t]
  \centering
  \includegraphics[width=1.0\textwidth]{diagram}
  \caption{Evaluation of a software modularization recovery algorithm}
  \label{fig:diagram}
  \end{figure*}

Unfortunately there are few systems with known reference modularizations and,
because it is costly to obtain reference modularizations, there are few
empirical studies, and most of them consider a couple of small and medium
systems.

We therefore propose to use synthetic, i.e., computer-generated, software
dependency networks, to evaluate software modularization recovery algorithms.
These networks are generated by parametrizable models and have an embedded
reference modularizations. The goal of an algorithm is, thus, to find
modularizations that are similar to the reference modularization embedded in the
network. With this approach we can CONTAR COM a large volume of test data that
is composed of networks of different sizes and controllable characteristics.

Of course the success of this approach depends on the similarity between
synthetic networks and networks extracted from real software systems. In this
paper we study three models and show that all of them, by means of a careful
choice of parameter values, are capable of producing software-like networks.

@@

Define: Component. Component dependency network. Directed unweighted graph.

Clustering based on dependencies.

Network theory research studies general properties of many types of networks by
using statistical analysis. In the last decade, it has been found that many
networks arising from sociology, biology, technology and other domains present
remarkable structural similarities. It has been shown that, in these networks,
the degree distribution is a power law: the number of vertices connected to k
edges, N(k), is proportional to $k^{\-gamma}$, where $\gamma$ is a positive
constant. These networks are called scale-free networks.

@@

Many models were proposed to explain the formation of scale-free networks. These
models are simple algorithms that can be proven, either formally or empirically,
to synthesize networks that are scale-free. Most models, though, do not account
for modular structure, a feature present in some networks, in particular
software networks. In this section we present three models that generate
directed networks with built-in modular decomposition: BCR+, CGW and LR.

@@

\begin{itemize}
\item Number of vertices, n;
\item Number of modules, m;
\item Probabilities p1, p2, p3, p4, summing 1;
\item Natural numbers e1, e2, e3, e4;
\item Number alpha, with $alpha \ge -1$
\end{itemize}

@@

% Word adjacency network: word Y follows word X => X->Y
\begin{figure}[!t]
\center
\includegraphics{tcp}
\caption{Triad concentration profiles (TCP) for two networks. On the left,
network extracted from the software system JabRef (see Appendix A). On the
right, word adjacency network for the Japanese language \cite{Milo2004}.}
\label{fig:profiles}
\end{figure}

@@

To support the evaluation of the metric, we have collected 131 networks. The
networks are described in detail in Appendix A.

@@

\appendix{Network Data Set}

>>>>

One of the most obvious ways to reduce the time needed to develop or maintain a
software system is to assign more developers to it. When combined with poor
management, though, adding developers may in fact raise the development time,
due to the increasing communication cost between developers \cite{Brooks1995}.
Poor management may result in developers working on overlapping tasks, writing
duplicated code, and inadvertently breaking the code written by another
developer.

%The development and maintenance of large software systems is a big challenge for
%software engineering \cite{Northrop2006}. In order to reduce the development
%time of such a system, a tempting idea is to assign more developers for the
%project.

In order to take advantage of a team of developers, therefore, a key issue is
the ability to decompose a system into weakly coupled modules, so that each
module is maintained by a distinct group of developers \cite{Parnas1972}.
Because of the weak coupling between modules, the need for communication between
groups of developers is reduced, and so is the development time.

The ability of decomposing a system into modules depends on global knowledge
about how its implementation components (e.g., functions, classes, variables)
interact. Often, specially in legacy systems, this information is not readily
available. Instead, it must be extracted from the system's source code. Even
then the task of finding modules may be overwhelming because of the large number
of components and interactions.

The task can be made easier by the use of software clustering algorithms, also
known as also known as architecture recovery algorithms \cite{Pollet2007}. They
automatically decompose a software system into weakly coupled modules by
analyzing the network of dependencies between its components.

% sw clustering is a task of reverse engineering \cite{Tonella2007}.

An important question for the adoption of a software clustering algorithm is
whether the decomposition it finds for a system is similar to a reference
decomposition for that system, i.e., a decomposition found by a group of
experienced developers. To answer this question one needs to test the algorithm
by applying it to a sample of systems with known reference decompositions, and
then compare the two decompositions for each system by means of a metric such as
MoJo \cite{Tzerpos1999} and EdgeSim \cite{Mitchell2001}.

Unfortunately there are few systems with known reference modularizations.
Furthermore, because it is costly to obtain reference modularizations
\cite{Koschke2000}, there are few empirical studies about software clustering
algorithms, most of them based on a couple of small and medium systems.

%  \begin{figure*}[!t]
%  \centering
%  \includegraphics[width=1.0\textwidth]{diagram}
%  \caption{Evaluation of a software clustering recovery algorithm}
%  \label{fig:diagram}
%  \end{figure*}

In this research we propose to evaluate software clustering algorithms by
applying them to synthetic, i.e., computer generated, component dependency
networks with built-in decompositions. With this approach it is feasible to test
algorithms with arbitrarily large samples in a controlled manner.

Of course, it is desirable that the synthetic networks resemble networks
extracted from real software systems. Hence, in this paper, we describe a study
about three models that synthesize networks with built-in decompositions. We
show empirically that all three models are capable of synthesizing networks that
resemble software networks.

@@

The implementation of a software system can be viewed as a network of
interacting components. In a object-oriented system, for example, classes
interact with other classes through mechanisms such as inheritance and
aggregation. By abstract the particular mechanisms of interaction, it is
possible to build a network of dependencies between components, represented as a
directed graph. This kind of network is the input for many software clustering
algorithms \cite{Mancoridis1998,Anquetil1999,Tzerpos2000,Andritsos2005}.

In this paper we will study networks of dependencies between classes in
object-oriented systems. We will call any such a network a \emph{software
network}.

%Define: Component. Component dependency network. Directed unweighted graph.
%Clustering based on dependencies.

Network theory research studies general properties of many types of networks by
using statistical analysis. In the last decade, it has been found that many
networks arising from sociology, biology, technology and other domains present
remarkable structural similarities. It has been shown that, in these networks,
the distribution of vertex degrees is a power law, i.e., the number of vertices
connected to $k$ edges, N($k$), is proportional to $k^{\-gamma}$, where $\gamma$
is a positive constant. These networks are called scale-free networks.

@@

Many models were proposed to explain the formation of scale-free networks. These
models are simple algorithms that can be proven, either formally or empirically,
to synthesize networks that are scale-free. Most models, though, do not account
for modular structure, a feature present in some networks, in particular
software networks. In this section we present three models that generate
directed networks with built-in modular decompositions: BCR+, CGW and LR.

@@

\begin{itemize}
\item Number of vertices, n;
\item Number of modules, m;
\item Probabilities p1, p2, p3, p4, summing 1;
\item Natural numbers e1, e2, e3, e4;
\item Constant $\alpha$, with $\alpha \ge -1$
\end{itemize}

@@

% Word adjacency network: word Y follows word X => X->Y
\begin{figure}[!t]
\center
\includegraphics{tcp}
\caption{Triad concentration profiles (TCP) for two networks. On the left,
network extracted from the software system JabRef (see the Appendix). On the
right, word adjacency network for the Japanese language \cite{Milo2004}.}
\label{fig:profiles}
\end{figure}

@@

To support the evaluation of the metric, we have collected 131 networks. The
networks are described in detail in the Appendix.

@@

\appendix %{Network Data Set}

This appendix lists the networks used in this work.

\subsection{Software Networks}
